\input{/Users/daniel/github/config/preamble.sty}%available at github.com/danimalabares/config
\input{/Users/daniel/github/config/thms-eng.sty}%available at github.com/danimalabares/config

\begin{document}

\begin{minipage}{\textwidth}
	\begin{minipage}{1\textwidth}
		Geometric Structures on Manifolds Seminar\hfill IMPA
		
		{\small\hfill\href{https://github.com/danimalabares/seminars}{github.com/danimalabares/seminars}}
		
		%{\small\hfill\href{https://github.com/Friday-seminar/}{github.com/Friday-seminar}}
	\end{minipage}
\end{minipage}\vspace{.2cm}\hrule

\vspace{10pt}

{\Huge Brody lemma and Kobayashi metric}

\hfill{\Large Misha Verbitsky}

\hfill{\large November 7, 2024}

\tableofcontents
\begin{upshot}\leavevmode
	Brody lemma says Kobayashi hyperbolic is equivalent to contaning $\mathbb{C}$.
\end{upshot}
$\mathsf{SO}(n)$ acts on $\mathbb{R}^n$ any $\mathsf{SO}(n)$-invariant quadratic form is proportional to $\sum x_i^2$. The orbits fo $\mathsf{SO}(n)$ are spheres…

\begin{defn}\leavevmode
	\textit{\textbf{Brody curve}} is a non-constant holomorphic map $f:\mathbb{C}\to M$ such that $|df|\leq C$ for some constant $C$.
\end{defn}

\begin{defn}\leavevmode
	$(\Delta_r,g_r)$ a disck of raduis $r$ in $\mathbb{C}$ and Poincaré metric rescaled ``in a such a way that the unit tangent vector to 0 has length 1". \textit{\textbf{Brody map}} is a holomorphic $\Delta_r \to M$ with $|df|\leq 1$ and $|df|_\partial=1$.
\end{defn}

\begin{thing4}{Lemma 1}\leavevmode
Let $f_r:\Delta_r \to M$ be a sequence of Brody maps, then $f_r$ converges uniformly to a Brody curve.
\end{thing4}

\begin{proof}\leavevmode
Took some time, using Kobayashi metric, theorems of uniform convergence.
\end{proof}

\begin{thm}[Brody lemma]\leavevmode
	Let $M$be a compact complex manifold that is not Kobayashi hyperbolic. Then $M$ contains a Brody curve.
\end{thm}

\begin{proof}\leavevmode
Somehow we have a sequence of maps, but they are not Brody. To have Brody curve we need to take the limit of Brody maps.
\end{proof}




\end{document}
